<!--
%\VignetteEngine{knitr::docco_linear}
%\VignetteIndexEntry{An Introduction to the corrplot package}
-->

An Introduction to **corrplot** Package
=======================================


```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(out.extra = 'style="display:block; margin: auto"',
               fig.align = "center", dev = "png")
```


Introduction
------------
The **corrplot** package is a graphical display of a correlation matrix, 
confidence interval. It also contains some algorithms to do matrix reordering. 
In addition, corrplot is good at details, including choosing color, text labels,
color labels, layout, etc.


Visualization methods
----------------------------
There are seven visualization methods (parameter `method`) in **corrplot** package, named `"circle"`, `"square"`, `"ellipse"`, `"number"`, `"shade"`, `"color"`, `"pie"`. 
```{r methods}
library(corrplot)
M <- cor(mtcars)
corrplot(M, method = "circle")
corrplot(M, method = "square")
corrplot(M, method = "ellipse")
corrplot(M, method = "number")
corrplot(M, method = "shade")
corrplot(M, method = "color")
corrplot(M, method = "pie")
```


Layout
-----------------------------
There are three layout types (parameter `type`), named `"full"` (default),
`"upper"` or `"lower"`, display full matrix, lower triangular or upper
triangular matrix.

```{r layout}
corrplot(M, type = "upper")
corrplot(M, type = "lower")
```

`corrplot.mixed()` is a wrapped function for mixed visualization style.
```{r mixed}
corrplot.mixed(M)
corrplot.mixed(M, lower = "ellipse", upper = "circle")
corrplot.mixed(M, lower = "square", upper = "circle")

```


Reorder a correlation matrix
----------------------------
Matrix reorder is very important for mining the hiden structure and pattern in
the matrix. There are four methods in corrplot (parameter `order`), named
`"AOE"`, `"FPC"`, `"hclust"`, `"alphabet"`.  More algorithms can be found in
[seriation](cran.r-project.org/package=seriation) package.

You can also reorder the matrix "manually" via function `corrMatOrder()`.

  - `"AOE"` is for the angular order of the eigenvectors. It is calculated from 
    the order of the angles $a_i$,
    
    $$
    a_i = 
    \begin{cases}
    			\tan (e_{i2}/e_{i1}), & \text{if $e_{i1}>0$;}
    			 \newline
    			\tan (e_{i2}/e_{i1}) + \pi, & \text{otherwise.}
    \end{cases}			
    $$
    
    where $e_1$ and $e_2$ are the largest two eigenvalues of the correlation 
    matrix.
    See [Michael Friendly (2002)](http://www.datavis.ca/papers/corrgram.pdf)
    for details.
  
  - `"FPC"` for the first principal component order.
  
  - `"hclust"` for hierarchical clustering order, and `"hclust.method"` for the
    agglomeration method to be used. `"hclust.method"` should be one of
    `"ward"`, `"single"`, `"complete"`, `"average"`, `"mcquitty"`, `"median"` or
    `"centroid"`.

  - `"alphabet"` for alphabetical order.

```{r order}
corrplot(M, order = "AOE")
corrplot(M, order = "hclust")
corrplot(M, order = "FPC")
corrplot(M, order = "alphabet")
```

If using `"hclust"`, `corrplot()` can  draw rectangles around the chart of corrrlation matrix based on the results of  hierarchical clustering.

```{r rectangles}
corrplot(M, order = "hclust", addrect = 2)
corrplot(M, order = "hclust", addrect = 3)
```


Using different color spectra
-----------------------------
We can also specify the color system, `colorRampPalette()` is very convenient
for generating color spectrum.

```{r color}
col1 <- colorRampPalette(c("#7F0000", "red", "#FF7F00", "yellow", "white",
                           "cyan", "#007FFF", "blue", "#00007F"))
col2 <- colorRampPalette(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
                           "#4393C3", "#2166AC", "#053061"))
col3 <- colorRampPalette(c("red", "white", "blue"))	
col4 <- colorRampPalette(c("#7F0000","red","#FF7F00","yellow","#7FFF7F",
                           "cyan", "#007FFF", "blue","#00007F"))
whiteblack <- c("white","black")

## using these color spectra
corrplot(M, order = "hclust", addrect = 2, col = col1(100))
corrplot(M, order = "hclust", addrect = 2, col = col2(50))
corrplot(M, order = "hclust", addrect = 2, col = col3(20))
corrplot(M, order = "hclust", addrect = 2, col = col4(10))
corrplot(M, order = "hclust", addrect = 2, col = whiteblack, bg = "gold2")

## you can also use the standard color palettes (package `grDevices`)
corrplot(M, order = "hclust", addrect = 2, col = heat.colors(100))
corrplot(M, order = "hclust", addrect = 2, col = terrain.colors(100))
corrplot(M, order = "hclust", addrect = 2, col = cm.colors(100))
```


Color legend and text legend
----------------------------
Parameter `cl.*` is for color legend, and `tl.*` if for text legend.

Here are some examples.
```{r color-label}
## remove color legend and text legend 
corrplot(M, order = "AOE", cl.pos = "n", tl.pos = "n")  

## bottom  color legend, diagonal text legend, rotate text label
corrplot(M, order = "AOE", cl.pos = "b", tl.pos = "d", tl.srt = 60)

## a wider color legend with numbers right aligned
corrplot(M, order = "AOE", cl.ratio = 0.2, cl.align = "r")
```


Dealing with a non-correlation matrix
-------------------------------------
```{r non-corr}
corrplot(abs(M),order = "AOE", col = col3(200), cl.lim = c(0,1))
## visualize a  matrix in [-100, 100]
ran <- round(matrix(runif(225, -100,100), 15))
corrplot(ran, is.corr = FALSE, method = "square")
## a beautiful color legend 
corrplot(ran, is.corr = FALSE, method = "ellipse", cl.lim = c(-100, 100))
```

If your matrix is rectangular, you can adjust the aspect ratio with the
`win.asp` parameter to make the matrix rendered as a square.
```{r non-corr-asp}
ran <- matrix(rnorm(70), ncol = 7)
corrplot(ran, is.corr = F, win.asp = .7, method = "circle")
```

Dealing with missing (NA) values
--------------------------------
By default, **corrplot** renders NA values as `"?"` characters. Using `na.label`
parameter, it is possible to use a different value (max. two characters are
supported).

```{r NAs}
M2 <- M
diag(M2) = NA
corrplot(M2)
corrplot(M2, na.label = "o")
corrplot(M2, na.label = "NA")
```


Using "plotmath" expressions in labels
--------------------------------------
Since version `0.78`, it is possible to use
[plotmath](https://artax.karlin.mff.cuni.cz/r-help/library/grDevices/html/plotmath.html)
expression in variable names. To activate plotmath rendering, prefix your label
with one of the characters `":"`, `"="` or `"$"`.

```{r plotmath}
M2 <- M[1:5,1:5]
colnames(M2) <- c("alpha", "beta", ":alpha+beta", ":a[0]", "=a[beta]")
rownames(M2) <- c("alpha", "beta", NA, "$a[0]", "$ a[beta]")
corrplot(M2)
```


Combine with the significance test
----------------------------------
```{r test}
cor.mtest <- function(mat, conf.level = 0.95){
  mat <- as.matrix(mat)
	n <- ncol(mat)
	p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
	diag(p.mat) <- 0
	diag(lowCI.mat) <- diag(uppCI.mat) <- 1
	for (i in 1:(n - 1)) {
		for (j in (i + 1):n) {
			tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
			p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
			lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
			uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
		}
	}
	return(list(p.mat, lowCI.mat, uppCI.mat))
}

res1 <- cor.mtest(mtcars,0.95)
res2 <- cor.mtest(mtcars,0.99)

## specialized the insignificant value according to the significant level
corrplot(M, p.mat = res1[[1]], sig.level = 0.2)
corrplot(M, p.mat = res1[[1]], sig.level = 0.05)
corrplot(M, p.mat = res1[[1]], sig.level = 0.01)

## leave blank on no significant coefficient
corrplot(M, p.mat = res1[[1]], insig = "blank")

## add p-values on no significant coefficient
corrplot(M, p.mat = res1[[1]], insig = "p-value")

## add all p-values
corrplot(M, p.mat = res1[[1]], insig = "p-value", sig.level = -1)

## add cross on no significant coefficient 
corrplot(M, p.mat = res1[[1]], order = "hclust", insig = "pch", addrect = 3)
```


Visualize confidence interval
-----------------------------
```{r ci}
## plot confidence interval(0.95, 0.95, 0.99), "rect" method
corrplot(M, low = res1[[2]], upp = res1[[3]], order = "hclust",
         rect.col = "navy", plotC = "rect",cl.pos = "n")
corrplot(M, p.mat = res1[[1]], low = res1[[2]], upp = res1[[3]],
         order = "hclust", pch.col = "red", sig.level = 0.01,
         addrect = 3, rect.col = "navy", plotC = "rect",cl.pos = "n")
```

