---
title: 'An Introduction to corrplot Package'
author: 'Taiyun Wei, Viliam Simko'
date: '`r Sys.Date()`'
output:
  prettydoc::html_pretty:
    theme: cayman
    toc: true
toc-title: 'Table of Contents'
vignette: >
  %\VignetteIndexEntry{An Introduction to corrplot Package}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align = 'center',
  fig.path = 'webimg/',
  fig.width = 6,
  fig.height = 6,
  dev = 'png')

get_os = function(){
  sysinf = Sys.info()
  if (!is.null(sysinf)){
    os = sysinf['sysname']
    if (os == 'Darwin')
      os = 'osx'
  } else { ## mystery machine
    os = .Platform$OS.type
    if (grepl('^darwin', R.version$os))
      os = 'osx'
    if (grepl('linux-gnu', R.version$os))
      os = 'linux'
  }
  tolower(os)
}
if(get_os() =='windows' & capabilities('cairo')){
  knitr::opts_chunk$set(dev.args = list(type='cairo'))
}
```

## Introduction

**corrplot** package is a graphical display of correlation matrix,
p-value and confidence interval. It also contains some algorithms to do
matrix reordering. In addition, corrplot is good at details, including
choosing color, text labels, color labels, layout, adding rectangles,
etc.

`corrplot()` have about 50 parameters, however the most common using are
only a few. In most scenes, you can get a correlation matrix plot with
only one or a few lines of code.

The mostly using parameters are `method`, `type`, `order` and `diag`.

There are seven visualization methods (parameter `method`) in
**corrplot** package, named `'circle'`, `'square'`, `'ellipse'`,
`'number'`, `'shade'`, `'color'`, `'pie'`. Color intensity of the glyph
is proportional to the correlation coefficients by default. For method
`'circle'` and `'square'`, the **areas** of circles or squares show the
absolute value of corresponding correlation coefficients. For method
`'number'`, the correlation coefficients are showed with corresponding
color on the graph. Method `'pie'` and `'shade'` came from Michael
Friendly's job (with some adjustment about the shade added on). and
`'ellipse'` came from D.J. Murdoch and E.D. Chow's job, see in section
References.

`corrplot.mixed()` is a wrapped function for mixed visualization style,
which can set the visual methods of lower and upper triangular
separately.

There are three layout types (parameter `type`): `'full'`, `'upper'` and
`'lower'`.

The correlation matrix can be reordered according to the correlation
matrix coefficients. This is important to identify the hidden structure
and pattern in the matrix.

```{r intro}
library(corrplot)
M = cor(mtcars)
corrplot(M, method = 'number') 
corrplot(M, method = 'color', order = 'alphabet') # colorful number
corrplot(M) # by default, method = 'circle'
corrplot(M, order = 'AOE') # after 'AOE' reorder
corrplot(M, method = 'shade', order = 'AOE', diag = FALSE) 
corrplot(M, method = 'square', order = 'FPC', type = 'lower', diag = FALSE)
corrplot(M, method = 'ellipse', order = 'AOE', type = 'upper')
corrplot.mixed(M, order = 'AOE')
corrplot.mixed(M, lower = 'shade', upper = 'pie', order = 'hclust')
```

## Reorder a correlation matrix

The details of four `order` algorithms, named `'AOE'`, `'FPC'`,
`'hclust'`, `'alphabet'` are as following.

-   `'AOE'` is for the angular order of the eigenvectors. It is
    calculated from the order of the angles $a_i$,

    $$
    a_i = 
    \begin{cases}
                \arctan (e_{i2}/e_{i1}), & \text{if $e_{i1}>0$;}
                 \newline
                \arctan (e_{i2}/e_{i1}) + \pi, & \text{otherwise.}
    \end{cases}         
    $$

    where $e_1$ and $e_2$ are the largest two eigenvalues of the
    correlation matrix. See [Michael Friendly
    (2002)](http://www.datavis.ca/papers/corrgram.pdf) for details.

-   `'FPC'` for the first principal component order.

-   `'hclust'` for hierarchical clustering order, and `'hclust.method'`
    for the agglomeration method to be used. `'hclust.method'` should be
    one of `'ward'`, `'single'`, `'complete'`, `'average'`,
    `'mcquitty'`, `'median'` or `'centroid'`.

-   `'alphabet'` for alphabetical order.

You can also reorder the matrix 'manually' via function
`corrMatOrder()`.

If using `'hclust'`, `corrplot()` can draw rectangles around the plot of
correlation matrix based on the results of hierarchical clustering.

```{r hclust}
corrplot(M, order = 'hclust', addrect = 2)
corrplot(M, method = 'square', diag = FALSE, order = 'hclust', 
         addrect = 3, rect.col = 'blue', rect.lwd = 3, tl.pos = 'd')
```

`corrRect()` can add rectangles on the plot with three ways(parameter
`index`, `name` and `namesMat`) after `corrplot()`.

Since R 4.1.0, we can use pipe operator `|>` very conveniently.

```{r rectangles}
# use namesMat parameter
r = rbind(c('gear', 'wt', 'qsec', 'carb'),
        c('wt', 'gear', 'carb', 'qsec'))
# Since R 4.1.0, the better way is the following one line code: 
# corrplot(M, order = 'AOE') |> corrRect(namesMat = r)
corrplot(M, order = 'AOE') -> p
corrRect(p, namesMat = r)

# show the return list of corrplot()
attributes(p)
print(p$corr, digits=2)
head(p$corrPos)
p$arg

# use name parameter
# Since R 4.1.0, the better way is the following one line code: 
# corrplot(M, order = 'AOE') |> corrRect(name = c('gear', 'wt', 'carb'))
corrplot(M, order = 'AOE') -> p
corrRect(p, name = c('gear', 'wt', 'carb'))
```

## Change color spectra, color-legend and text-legend

The color of the correlation plots can be customized by `col`. The
function `colorRampPalette()` is very convenient for generating color
spectrum. Other option would be to use `RcolorBrewer` package.

Parameters group `cl.*` is for color-legend. The common-using are:

-   `cl.pos` is for the position of color labels. It is character or
    logical. If character, it must be one of `'r'` (means right, default
    if `type='upper'` or `'full'`), `'b'` (means bottom, default if
    `type='lower'`) or `'n'`(means don't draw color-label).
-   `cl.ratio` is to justify the width of color-legend, 0.1\~0.2 is
    suggested.

Parameters group `tl.*` is for text-legend. The common-using are:

-   `tl.pos` is for the position of text labels. It is character or
    logical. If character, it must be one of `'lt'`, `'ld'`, `'td'`,
    `'d'` or `'n'`. `'lt'`(default if `type='full'`) means left and top,
    `'ld'`(default if `type='lower'`) means left and diagonal,
    `'td'`(default if `type='upper'`) means top and diagonal(near),
    `'d'` means diagonal, `'n'` means don't add text-label.
-   `tl.cex` is for the size of text label (variable names).
-   `tl.srt` is for text label string rotation in degrees.

```{r color}
## diverging color example, suitable for matrix in [-N, N]
col1 = colorRampPalette(c('#7F0000', 'red', '#FF7F00', 'yellow', 'white',
                           'cyan', '#007FFF', 'blue', '#00007F'))
col2 = colorRampPalette(c('red', 'white', 'blue'))	

library(RColorBrewer)
# use RcolorBrewer color
corrplot(M, type = 'upper', order = 'hclust', 
         col = brewer.pal(n = 8, name = 'RdBu'))

## bottom color legend, diagonal text legend, rotate text label
corrplot(M, order = 'AOE', cl.pos = 'b', tl.pos = 'd', 
         col = col1(10), diag = FALSE)

## text labels rotated 45 degrees and  wider color legend with numbers right aligned
corrplot(M, type = 'lower', order = 'hclust', tl.col = 'black', 
         cl.ratio = 0.2, tl.srt = 45, col = col2(10))

## remove color legend, text legend and principal diagonal glyph
corrplot(M, order = 'AOE', cl.pos = 'n', tl.pos = 'n', 
         col = c('white', 'black'), bg = 'gold2')  
```

## Visualize non-correlation matrix, NA value and math label

We can visualize a non-correlation matrix by set `is.corr=FALSE`, and
assign colors by `col.lim`. If the matrix have both positive and
negative values, the matrix transformation keep every values
positiveness and negativeness.

If your matrix is rectangular, you can adjust the aspect ratio with the
`win.asp` parameter to make the matrix rendered as a square.

```{r non-corr}
## color example, suitable for matrix in [-N, 0] or [0, N]
col3 = hcl.colors(10, "YlOrRd", rev = TRUE)

N = matrix(runif(80, 20, 30), 8)
corrplot(N, is.corr = FALSE, col.lim = c(20, 30), 
         col = col3, cl.pos = 'b', win.asp = 0.8)
```

Notice: when `is.corr=TRUE`, `col.lim` only affect the color legend If
you change it, the color on correlation matrix plot is still assigned on
`c(-1, 1)`

```{r col-lim}
# when is.corr=TRUE, col.lim only affect the color legend display
corrplot(M/2)
corrplot(M/2, col.lim=c(-0.5, 0.5))
```

By default, **corrplot** renders NA values as `'?'` characters. Using
`na.label` parameter, it is possible to use a different value (max. two
characters are supported).

Since version `0.78`, it is possible to use
[plotmath](https://www.rdocumentation.org/packages/grDevices/topics/plotmath)
expression in variable names. To activate plotmath rendering, prefix
your label with `'$'`.

```{r NA-math}
M2 = M
diag(M2) = NA
colnames(M2) = rep(c('$alpha+beta', '$alpha[0]', '$alpha[beta]'), 
                   c(4, 4, 3))
rownames(M2) = rep(c('$Sigma[i]^n', '$sigma',  '$alpha[0]^100', '$alpha[beta]'), 
                   c(2, 4, 2, 3))
corrplot(10*abs(M2), col = col3, is.corr = FALSE, 
         col.lim = c(0, 10), tl.cex = 1.5)
```

## Visualize p-value and confidence interval

`corrplot()` can also visualize p-value and confidence interval on the
correlation matrix plot. Here are some important parameters.

About p-value:

-   `p.mat` is the p-value matrix, if `NULL`, parameter `sig.level`,
    `insig, pch`, `pch.col`, `pch.cex` are invalid.
-   `sig.level` is significant level, with default value 0.05. If the
    p-value in `p-mat` is bigger than `sig.level`, then the
    corresponding correlation coefficient is regarded as insignificant.
    If `insig` is `'label_sig'`, `sig.level` can be an increasing vector
    of significance levels, in which case `pch` will be used once for
    the highest p-value interval and multiple times (e.g. `'*'`, `'**'`,
    `'***'`) for each lower p-value interval.
-   `insig` Character, specialized insignificant correlation
    coefficients, `'pch'` (default), `'p-value'`, `'blank',` `'n'`, or
    `'label_sig'`. If `'blank'`, wipe away the corresponding glyphs; if
    `'p-value'`, add p-values the corresponding glyphs; if `'pch'`, add
    characters (see pch for details) on corresponding glyphs; if `'n'`,
    don't take any measures; if `'label_sig'`, mark significant
    correlations with `pch` (see `sig.level`).
-   `pch` is for adding character on the glyphs of insignificant
    correlation coefficients (only valid when insig is `'pch'`). See
    `?par` .

About confidence interval:

-   `plotCI` is character for the method of plotting confidence
    interval. If `'n'`, don't plot confidence interval. If `'rect'`,
    plot rectangles whose upper side means upper bound and lower side
    means lower bound respectively.
-   `lowCI.mat` is the matrix of the lower bound of confidence interval.
-   `uppCI.mat` is the Matrix of the upper bound of confidence interval.

We can get p-value matrix and confidence intervals matrix by
`cor.mtest()` which returns a list containing:

-   `p` is the p-values matrix.
-   `lowCI` is the lower bound of confidence interval matrix.
-   `uppCI` is the lower bound of confidence interval matrix.

```{r test}
testRes = cor.mtest(mtcars, conf.level = 0.95)

## specialized the insignificant value according to the significant level
corrplot(M, p.mat = testRes$p, sig.level = 0.05, order = 'hclust', addrect = 2) 

## leave blank on no significant coefficient
corrplot(M, p.mat = testRes$p, method = 'circle', type = 'lower', insig='blank',
         addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

## add p-values on no significant coefficients
corrplot(M, p.mat = testRes$p, insig = 'p-value')

## add all p-values
corrplot(M, p.mat = testRes$p, insig = 'p-value', sig.level = -1)

## add significant level stars
corrplot(M, p.mat = testRes$p, method = 'color', diag = FALSE, type = 'upper',
         sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.9, 
         insig = 'label_sig', pch.col = 'grey20', order = 'AOE')

## add significant level stars and cluster rectangles
corrplot(M, p.mat = testRes$p, tl.pos = 'd', order = 'hclust', addrect = 2,
         insig = 'label_sig', sig.level = c(0.001, 0.01, 0.05), 
         pch.cex = 0.9, pch.col = 'grey20')
```

Visualize confidence interval.

```{r confidence-interval}
# Visualize confidence interval
corrplot(M, lowCI = testRes$lowCI, uppCI = testRes$uppCI, order = 'hclust',
         tl.pos = 'd', rect.col = 'navy', plotC = 'rect', cl.pos = 'n')

# Visualize confidence interval and cross the significant coefficients
corrplot(M, p.mat = testRes$p, lowCI = testRes$lowCI, uppCI = testRes$uppCI,
         addrect = 3, rect.col = 'navy', plotC = 'rect', cl.pos = 'n')
```

## References

Michael Friendly (2002). Corrgrams: Exploratory displays for correlation
matrices. The American Statistician, 56, 316--324.

D.J. Murdoch, E.D. Chow (1996). A graphical display of large correlation
matrices. The American Statistician, 50, 178--180.
